Dynamic Witnesses for Static Type Errors

### 2015-11-20 paper/pldi16-submission/overview.tex
remove sub-term edges from the graph
tweak

\paragraph{Problem: How many inputs does a function take?}
%
There is another wrinkle though, how did we know
that @fac@ takes a single argument instead of two
(or none)?
%
It is clear, syntactically, that @fac@ takes \emph{at least} one
argument, but in a higher-order language with currying, syntax can be
deceiving.
%
Consider the following definition:
%
\begin{code}
  let incAllByOne = List.map (+ 1)
\end{code}
%
Is @incAllByOne@ a function?
%
If so, how many arguments does it take?
%
The \ocaml\ compiler deduces that @incAllByOne@ takes a single argument
because @List.map@'s \emph{type} says it takes two arguments, and it is
partially applied to @(+ 1)@.
%
As we are dealing with ill-typed programs we do not have the luxury of
typing information.

@@

For example, evaluating the expression @1+2+3@ would produce the
graph in Figure~\ref{fig:simple-reduction-hi}.
%
\begin{figure}[t]
  \centering
  \includegraphics[width=\linewidth]{simple.pdf}
  \caption{The reduction graph for \texttt{1+2+3}. The two edges
    produced by the transition from \texttt{1+2+3} to \hbox{\texttt{3+3}}
    are highlighted.}
\label{fig:simple-reduction-hi}
\end{figure}
%
Notice that when we transition from @1+2+3@ to @3+3@ we collect
both that edge \emph{and} an edge from the sub-term @1+2@ to @3@.
%
These additional edges allow us to implement two common debugging
operations \emph{post-hoc}: ``step into'' to zoom in on a specific
function call, and ``step over'' to skip over an uninteresting
sub-computation.

>>>>

\paragraph{Problem: How many inputs does a function take?}
%
There is another wrinkle though, how did we know
that @fac@ takes a single argument instead of two
(or none)?
%
It is clear, syntactically, that @fac@ takes \emph{at least} one
argument, but in a higher-order language with currying, syntax can be
deceiving.
%
Consider the following definition:
%
\begin{code}
  let incAllByOne = List.map (+ 1)
\end{code}
%
Is @incAllByOne@ a function?
%
If so, how many arguments does it take?
%
The \ocaml\ compiler deduces that @incAllByOne@ takes a single argument
because the \emph{type} of \hbox{@List.map@'s} says it takes two arguments, and it is
partially applied to @(+ 1)@.
%
As we are dealing with ill-typed programs we do not have the luxury of
typing information.

@@

For example, evaluating the expression @1+2+3@ would produce the
graph in Figure~\ref{fig:simple-reduction-hi}.
%
\begin{figure}[t]
  \centering
  \includegraphics[height=2in]{simple.pdf}
  \caption{The reduction graph for \texttt{1+2+3}. The two edges
    produced by the transition from \texttt{1+2+3} to \hbox{\texttt{3+3}}
    are highlighted.}
\label{fig:simple-reduction-hi}
\end{figure}
%
Notice that when we transition from @1+2+3@ to @3+3@ we collect
both that edge \emph{and} an edge from the sub-term @1+2@ to @3@.
%
These additional edges allow us to implement two common debugging
operations \emph{post-hoc}: ``step into'' to zoom in on a specific
function call, and ``step over'' to skip over an uninteresting
sub-computation.
